FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2018-06-11 ECOOP18/sections/Discussion.tex
spellcheck done

\yanlin{Bruno: discussed multiple overriding and the two designs}
One non-orthogonal extension to \MIM{} could be to generalize the
model to allow multiple hierarchical method overriding, meaning that,
we allow overriding methods to update multiple branches instead of
only one branch. This feature offers a more fine-grained mechanism for merging, and can be helpful to easily understand the structure of the hierarchy. Multiple overriding would be useful in the following situation, for example:

@@

With this approach, branches \lstinline|A| and \lstinline|B| are merged in the sense that they share the same code, which can be separately updated in future interfaces. Another approach would be to deeply merge the branches, with similar effect as introducing an intermediate interface \lstinline|AB| to explicitly merge the two branches. However, this approach is problematic because there is no clear mechanism of identifying and further updating the merged branches. This could be an interesting future work to explore.

>>>>

\yanlin{Bruno: discussed multiple overriding and the two designs}
One non-orthogonal extension to \MIM{} could be to generalize the
model to allow multiple hierarchical method overriding, meaning that,
we allow overriding methods to update multiple branches instead of
only one branch. This feature offers a more fine-grained mechanism for merging and can be helpful to easily understand the structure of the hierarchy. Multiple overriding would be useful in the following situation, for example:

@@

With this approach, branches \lstinline|A| and \lstinline|B| are merged in the sense that they share the same code, which can be separately updated in future interfaces. Another approach would be to deeply merge the branches, with similar effect as introducing an intermediate interface \lstinline|AB| to explicitly merge the two branches. However, this approach is problematic because there is no clear mechanism for identifying and further updating the merged branches. This could be an interesting future work to explore.

### 2018-06-11 ECOOP18/sections/RelatedWork.tex
remove line number, comments, compress to 25pages.
spellcheck done

Although C++ supports hierarchical dispatching, it does not support
hierarchical overriding.  However, there are some possible
workarounds that can mimick hierarchical
overriding, including the \emph{MiddleMan} approach\footnote{\url{https://stackoverflow.com/questions/44632250/can-i-do-mimic-things-likes-this-partial-ove}\\\url{rride-in-c}}, the \emph{interface classes} pattern as described in Section 25.6 of~\cite{stroustrup1995}, the \emph{LotterySimulation} discussion in~\cite{stroustrup1994design}. Since these workarounds share the same spirit, we will discuss in detail the \emph{MiddleMan} approach, with the code shown in 
Figure~\ref{fig:middleman}. In this example, classes \texttt{A} and
\texttt{B} are two classes that both define a method with the same
name \texttt{m}
unintentionally.

Class \texttt{MiddleMan}, as its name suggests, acts as a middle man between its class \texttt{C} and its parents \texttt{A, B}. \texttt{MiddleMan} defines a virtual method \texttt{m} that overrides a parent method \texttt{m} and delegates the implementation to another method \texttt{m\_impl} that takes $\kwthis$ as a parameter. C++ supports method overloading, so that multiple \texttt{m\_impl} methods with different parameter types can coexist. When defining class \texttt{C}, we specify the parents to be \texttt{MiddleMan<A>, MiddleMan<B>} instead of \texttt{A, B}. In this way, programmers may define new versions of \texttt{A.m} and \texttt{B.m} in class \texttt{C} by providing the corresponding \texttt{m\_impl} methods. Then in the client code, the method call \texttt{((A*)c)->m()} will print out string \texttt{"MA2"}, as expected. Although this workaround can help us defining partial method overrides to a certain extent, the drawbacks are obvious. Firstly, the approach is complex and requires the programmer to fully understand this approach. Moreover, the lack of direct syntax support makes MiddleMan code cumbersome to write. Finally, the approach is ad-hoc, meaning that the class \texttt{MiddleMan} shown in Figure~\ref{fig:middleman} is not general enough to be used in other cases: more middle-men are needed if partial method overrides happen in other classes; and it is even worse when return types differ.

@@

\noindent  {\bf Languages using hygienicity.}
In NextGen/MixGen~\cite{Allen:2003:FAG:949305.949316},
HygJava~\cite{kusmierek2007hygienic} and
Magda~\cite{Bono:2012:MNL:2367163.2367198}, \emph{hygienicity} is
proposed to deal with unintentional method conflicts. The idea is 
to give a method a unique identifier by prefixing the name
with an unambiguous path. As shown in Figure~\ref{fig:magda_prefix},
the prefix \lstinline|HelloWorld| in method call \lstinline|(new HelloWorld []).HelloWorld.MainMatter()| is mandatory. So writing programs in these languages is
tedious if not supported by a specialized IDE, that aids filling
prefix/method information.  The advantage of this
approach, compared to ours, is that it does not require any additional
notion for method dispatching. Indeed
the compilation strategy is simple, just by generating conventional code (say
in Java or C++) with
method names attached with prefixes. Unfortunately, the disadvantage
is that some expressive power is lost. In particular \emph{merging}
methods arising from diamond inheritance is not possible because the 
methods have different prefixes. 
As shown in Figure~\ref{fig:magda_code}, two methods
\lstinline|m| from different branches \lstinline|A| and \lstinline|B|
cannot be overridden by the method \lstinline|m| in \lstinline|C|
because they are regarded as unrelated methods, and \lstinline|m| in
\lstinline|C| is just another new method that has nothing to do with
\lstinline|A.m| or \lstinline|B.m|. The reason is that in these
hygienic approaches, path names are used to distinguish different
methods. In contrast, our model can deal with unintentional conflicts,
as well as merged methods because our semantics is not simply based on
prefixing. Instead, our model keeps the names of methods unchanged,
and our direct operational semantics takes static and
dynamic type information into account at runtime when doing method
dispatching. Finally, the multiple inheritance model in Magda is based on
Mixins, whereas \MIM{} is based on traits. Thus, Magda inherits all
limitations of Mixins (such as the linearization problem, etc).

\begin{figure*}[t]
	\saveSpaceFig
	\begin{lstlisting}
mixin HelloWorld of Object =
	new Object MainMatter()
	begin
		"Hello world".String.print();
	end;
end;
(new HelloWorld []).HelloWorld.MainMatter();
	\end{lstlisting}
	\caption{Full-qualified name of method calls in Magda.}
	\label{fig:magda_prefix}
	\saveSpaceFig
\end{figure*}

\begin{figure*}[t]
	\saveSpaceFig
\begin{lstlisting}
mixin A of Object =
	new String m()
	begin
		return "A";
	end;
end; 
mixin B of Object = 
	new String m()
	begin
		return "B";
	end;
end;
mixin C of A, B = 
	new String m()
	begin
		return "C";
	end;
end;
\end{lstlisting}
	\caption{Code in Magda.}
	\label{fig:magda_code}
	\saveSpaceFig
\end{figure*}

>>>>

Although C++ supports hierarchical dispatching, it does not support
hierarchical overriding.  However, there are some possible
workarounds that can mimic hierarchical
overriding, including the \emph{MiddleMan} approach\footnote{\url{https://stackoverflow.com/questions/44632250/can-i-do-mimic-things-likes-this-partial-ove}\\\url{rride-in-c}}, the \emph{interface classes} pattern as described in Section 25.6 of~\cite{stroustrup1995}, the \emph{LotterySimulation} discussion in~\cite{stroustrup1994design}. Since these workarounds share the same spirit, we will discuss in detail the \emph{MiddleMan} approach, with the code shown in 
Figure~\ref{fig:middleman}. In this example, classes \texttt{A} and
\texttt{B} are two classes that both define a method with the same
name \texttt{m}
unintentionally.

Class \texttt{MiddleMan}, as its name suggests, acts as a middleman between its class \texttt{C} and its parents \texttt{A, B}. \texttt{MiddleMan} defines a virtual method \texttt{m} that overrides a parent method \texttt{m} and delegates the implementation to another method \texttt{m\_impl} that takes $\kwthis$ as a parameter. C++ supports method overloading, so that multiple \texttt{m\_impl} methods with different parameter types can coexist. When defining class \texttt{C}, we specify the parents to be \texttt{MiddleMan<A>, MiddleMan<B>} instead of \texttt{A, B}. In this way, programmers may define new versions of \texttt{A.m} and \texttt{B.m} in class \texttt{C} by providing the corresponding \texttt{m\_impl} methods. Then in the client code, the method call \texttt{((A*)c)->m()} will print out the string \texttt{"MA2"}, as expected. Although this workaround can help us defining partial method overrides to a certain extent, the drawbacks are obvious. Firstly, the approach is complex and requires the programmer to fully understand this approach. Moreover, the lack of direct syntax support makes MiddleMan code cumbersome to write. Finally, the approach is ad-hoc, meaning that the class \texttt{MiddleMan} shown in Figure~\ref{fig:middleman} is not general enough to be used in other cases: more middlemen are needed if partial method overrides happen in other classes; and it is even worse when return types differ.

@@

\noindent  {\bf Languages using hygienicity.}
In NextGen/MixGen~\cite{Allen:2003:FAG:949305.949316},
HygJava~\cite{kusmierek2007hygienic} and
Magda~\cite{Bono:2012:MNL:2367163.2367198}, \emph{hygienicity} is
proposed to deal with unintentional method conflicts. The idea is 
to give a method a unique identifier by prefixing the name
with an unambiguous path. As shown in Figure~\ref{fig:magda_prefix},
the prefix \lstinline|HelloWorld| in the method call \lstinline|(new HelloWorld []).HelloWorld.MainMatter()| is mandatory. So writing programs in these languages is
tedious if not supported by a specialized IDE, that aids filling
prefix/method information.  The advantage of this
approach, compared to ours, is that it does not require any additional
notion for method dispatching. Indeed
the compilation strategy is simple, just by generating conventional code (say
in Java or C++) with
method names attached with prefixes. Unfortunately, the disadvantage
is that some expressive power is lost. In particular \emph{merging}
methods arising from diamond inheritance is not possible because the 
methods have different prefixes. 
As shown in Figure~\ref{fig:magda_code}, two methods
\lstinline|m| from different branches \lstinline|A| and \lstinline|B|
cannot be overridden by the method \lstinline|m| in \lstinline|C|
because they are regarded as unrelated methods, and \lstinline|m| in
\lstinline|C| is just another new method that has nothing to do with
\lstinline|A.m| or \lstinline|B.m|. The reason is that in these
hygienic approaches, path names are used to distinguish different
methods. In contrast, our model can deal with unintentional conflicts,
as well as merged methods because our semantics is not simply based on
prefixing. Instead, our model keeps the names of methods unchanged,
and our direct operational semantics takes static and
dynamic type information into account at runtime when doing method
dispatching. Finally, the multiple inheritance model in Magda is based on
Mixins, whereas \MIM{} is based on traits. Thus, Magda inherits all
limitations of Mixins (such as the linearization problem, etc).

% \begin{figure*}[t]
% 	% \saveSpaceFig
% 	\begin{lstlisting}
% mixin HelloWorld of Object =
% 	new Object MainMatter()
% 	begin "Hello world".String.print();	end;
% end;
% (new HelloWorld []).HelloWorld.MainMatter();
% 	\end{lstlisting}
%     \saveSpaceFig
% 	\caption{Full-qualified name of method calls in Magda.}
% 	\label{fig:magda_prefix}
% 	% \saveSpaceFig
% \end{figure*}

\begin{figure*}[t]
    % \saveSpaceFig
    \begin{lstlisting}
mixin HelloWorld of Object =
    new Object MainMatter()
    begin 
        "Hello world".String.print(); 
    end;
end;
(new HelloWorld []).HelloWorld.MainMatter();
    \end{lstlisting}
    \saveSpaceFig
    \caption{Full-qualified name of method calls in Magda.}
    \label{fig:magda_prefix}
    % \saveSpaceFig
\end{figure*}

\begin{figure*}[t]
	% \saveSpaceFig
\begin{lstlisting}
mixin A of Object =
	new String m()
	begin
		return "A";
	end;
end; 
mixin B of Object = 
	new String m()
	begin
		return "B";
	end;
end;
mixin C of A, B = 
	new String m()
	begin
		return "C";
	end;
end;
\end{lstlisting}
    \saveSpaceFig
	\caption{Code in Magda.}
	\label{fig:magda_code}
\end{figure*}

% \begin{figure*}[t]
%     % \saveSpaceFig
% \begin{lstlisting}
% mixin A of Object =
%     new String m() begin return "A"; end;
% end; 
% mixin B of Object = 
%     new String m() begin return "B"; end;
% end;
% mixin C of A, B = 
%     new String m() begin return "C"; end;
% end;
% \end{lstlisting}
%     \saveSpaceFig
%     \caption{Code in Magda.}
%     \label{fig:magda_code}
%     % \saveSpaceFig
% \end{figure*}
